### 2023-10-05 RISE_Projektbeschreibung/RISE_Projektbeschreibung.tex
uni

If the temperature is below some critical temperature $T_c$ they can condense into a macroscopic occupation of the ground state and allow for the famous perfect conductivity.
The two prominent collective excitations in a superconductor are the so-called Higgs and phase modes \cite{higgs, bo22}.
The former is energetically located at the band edge and sometimes referred to as the amplitude mode.
The latter originates in the invariance of the complex phase. 
Changing the phase does not affect the energy, thus the mode is located at 0.

A well-known and often used model is the so-called Hubbard model. Its Hamiltonian reads
\begin{equation}
    \label{eqn:hubbard}
    H = -t \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma}  + U \sum_i c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger c_{i\downarrow} c_{i\uparrow}\,,
\end{equation}
where the $\langle i, j \rangle$ denotes the summation over next neighbours, $t$ is a hopping strength, 
$U$ denotes the on-site interaction strength and $c_{i\sigma}^{(\dagger)}$ annihilates (creates) an electron with spin $\sigma$ on lattice site $i$.
Most commonly, $U$ is positive and represents the repulsive Coulomb interaction of two electrons.
In this case, the model exhibits antiferromagnetism (AFM).

However, negative values for $U$ are also possible \cite{sentef}. They are oftentimes achieved by the aforementioned effective interactions mediated by phonons.
If the system is precisely at half-filling and the underlying lattice of the system is bipartite, e.g. a square or simple cubic lattice, 
the system can be transformed by a specific kind of particle-hole transformation \cite{micnas90}.
This transformation effectively flips the sign of $U$, allowing researches to use the previously known antiferromagnetic results.
However, instead of antiferromagnetism, the system now exhibits the coexistence of two distinct phases.
Namely, s-wave superconductivity as described by the aforementioned BCS-theory and a so-called charge-density wave (CDW).
The latter describes a system, where the electron density exhibits a wave like behaviour, similar to the spins in an antiferromagnetic phase.

@@

Handling a quartic Hamiltonian, such as \autoref{eqn:hubbard}, is practically very difficult up to impossible,
as the Hilbert spaces grow exponentially with the system size.
The mean field approximation aims to reduce those kinds of Hamiltonians to a bilinear form by neglecting fluctuations around the the equilibrium.
Therefore, one introduces so-called normal ordered operators
\begin{equation}
    :A: = A - \expec{A}_0\,.
\end{equation}
These allow rewriting the quartic term in the Hubbard Hamiltonian to
\begin{subequations}
\begin{align}
    c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger c_{i\downarrow} c_{i\uparrow} &= :c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger c_{i\downarrow} c_{i\uparrow}:\\
        &+\expec{c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger} :c_{i\downarrow} c_{i\uparrow}: 
            + :c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger: \expec{c_{i\downarrow} c_{i\uparrow}} \\
        &-\expec{c_{i\uparrow}^\dagger c_{i\downarrow}} :c_{i\downarrow}^\dagger c_{i\uparrow}: 
            - :c_{i\uparrow}^\dagger c_{i\downarrow}: \expec{c_{i\downarrow}^\dagger c_{i\uparrow}} \\
        &+\expec{c_{i\uparrow}^\dagger c_{i\uparrow}} :c_{i\downarrow}^\dagger c_{i\downarrow}: 
            + :c_{i\uparrow}^\dagger c_{i\uparrow}: \expec{c_{i\downarrow}^\dagger c_{i\downarrow}} + \text{const}\,,
\end{align}
\end{subequations}
where the first term $:c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger c_{i\downarrow} c_{i\uparrow}:$ describes only small fluctuations that are neglected.
The remainder is only bilinear in terms of operators as the expectation values are merely numbers.

The challenge here is obtaining these expectation values as they depend on the mean-field Hamiltonian, which in turn depends on the expectation values.
The solution to this is a self-consistency calculation, for which one selects some starting values,
solves the Hamiltonian for the new starting values and iterates until convergence is achieved.

Afterwards, we use the iterated-equations-of-motion method (iEoM) to analyse the system for occurring excitations \cite{philip18, Kalthoff2017}.
This method is based on the Heisenberg equations of motion 
\begin{equation}
    \timeDeriv A(t) = \frac{i}{\hbar} [H, A(t)]\,.
\end{equation}
To solve this equation, one needs an operator basis, which is complete in the sense, that each commutator $[H, A(t)]$ can be represented by this basis.
Practically, this is often requires an infinite basis, which is impossible to implement numerically.
The idea of the iEoM is, to iterate the commutation process and truncate at some point, assuming that the error becomes smaller with each iteration.

Furthermore, an operator scalar product $(A|B)$ is introduced, which allows reducing the equations of motion from the operator form above to
\begin{equation}
    \timeDeriv \mathbf{c}(t) = \mathcal{M} \mathbf{c}(t)\,,
\end{equation}
where $\mathbf{c}(t)$ is a vector of time-dependent complex \emph{numbers} and a matrix
\begin{equation}
    \mathcal{M}_{ij} := (A_i | [H, A_j])\,,
\end{equation}
where $A_i, A_j$ are the operators from the aforementioned operator basis.

All system dynamics are now contained within the complex-valued matrix $\mathcal{M}$, which can be analysed using a plethora of well-established numerical methods.

>>>>

\vspace{10pt}
\begin{minipage}{0.40\textwidth}
    If the temperature is below some critical temperature $T_c$ the Cooper pairs can condense into a macroscopic occupation of the lowest energy level 
    and allow for the famous perfect conductivity.
    The two prominent collective excitations in a superconductor are the so-called Higgs and phase modes \cite{higgs, bo22}.
    The former is energetically located at the band edge and sometimes referred to as the amplitude mode.
    The latter originates from the invariance of the complex phase. 
    Changing the phase does not affect the energy, thus the mode is located at 0 energy in neutral superfluids.
\end{minipage}
\hspace{0.01\textwidth}
\begin{minipage}{0.58\textwidth}
    \begin{center}
        \includegraphics[width=\textwidth]{images/resolvent_general.pdf}
        \mcaption{1}{Example results for the spectral density in the superconducting phase.
            The two lines show the phase mode at $z=0$ and the Higgs mode at the lower band edge.}
    \end{center}
\end{minipage}
\vspace{10pt}

A well-known and often used model is the so-called extended Hubbard model. Its Hamiltonian reads
\begin{equation}
    \label{eqn:hubbard}
    H = -t \sum_{\langle i, j \rangle, \sigma} c_{i\sigma}^\dagger c_{j\sigma} 
    + U \sum_i c_{i\uparrow}^\dagger c_{i\downarrow}^\dagger c_{i\downarrow} c_{i\uparrow}
    + V \sum_{\langle i, j \rangle, \sigma, \sigma'} n_{i\sigma}n_{j\sigma'}\,,
\end{equation}
where the $\langle i, j \rangle$ denotes the summation over next neighbours on a lattice, $t$ is a hopping strength, 
$U$ denotes the on-site and $V$ the inter-site interaction strength and $c_{i\sigma}^{(\dagger)}$ annihilates (creates) an electron with spin $\sigma$ on lattice site $i$.
The number operator $n_{i\sigma} = c_{i\sigma}^\dagger c_{i\sigma}$ counts the number of electrons with spin $\sigma$ on lattice site $i$.
Most commonly, $U$ is positive and represents the repulsive Coulomb interaction of two electrons.
In this case, the model exhibits antiferromagnetism (AFM), if $|V|$ is not too large.

However, negative values for $U$ are also possible \cite{sentef}. They are oftentimes achieved by the aforementioned effective interactions mediated by phonons.
If the system is precisely at half-filling and the underlying lattice of the system is bipartite, e.g. a square or simple cubic lattice, 
the system can be transformed by a specific kind of particle-hole transformation \cite{micnas90}.
This transformation effectively flips the sign of $U$, allowing researches to use the previously known antiferromagnetic results.
However, instead of antiferromagnetism, the system now exhibits the coexistence of two distinct phases.
Namely, s-wave superconductivity as described by the aforementioned BCS-theory and a so-called charge-density wave (CDW).
The latter describes a system, where the electron density exhibits a wave like behaviour, similar to the spins in an antiferromagnetic phase.
This degeneracy is broken by any finite $V$.
\newpage

@@

\begin{minipage}{0.40\textwidth}
    Handling a quartic Hamiltonian, such as \autoref{eqn:hubbard}, is practically very difficult or even impossible,
    as the Hilbert spaces grow exponentially with the system size.
    The mean field approximation aims to reduce those kinds of Hamiltonians to a bilinear form by neglecting fluctuations around the the equilibrium.
    The remainder of the Hamiltonian is a set of products of expectation values and bilinear operators.

    The challenge here is obtaining these expectation values as they depend on the mean-field Hamiltonian, which in turn depends on the expectation values.
    The solution to this is a self-consistency calculation, for which one selects some starting values,
    solves the Hamiltonian for the new starting values and iterates until convergence is achieved.
\end{minipage}
\hspace{0.01\textwidth}
\begin{minipage}{0.58\textwidth}
    \begin{center}
        \includegraphics[width=\textwidth]{images/phase_boundaries_T0.pdf}
        \mcaption{1}{Example phase diagram obtained by mean field calculations.}
    \end{center}
\end{minipage}
\vspace{10pt}

Afterwards, we use the iterated-equations-of-motion method (iEoM) to analyse the system for occurring excitations \cite{philip18, Kalthoff2017}.
This method is based on the Heisenberg equations of motion 
\begin{equation}
    \timeDeriv A(t) = \frac{i}{\hbar} [H, A(t)]\,.
\end{equation}
To solve this equation, one needs an operator basis, which is complete in the sense, that each commutator $[H, A(t)]$ can be represented by this basis.
Practically, this is often requires an infinite basis, which is impossible to implement numerically.
The idea of the iEoM is, to iterate the commutation process and truncate at some point, assuming that the error becomes smaller with each iteration.
Introducing an operator scalar product $(A | B)$ allows us to reduce the operator-valued differential equations to simple matrices.
All system dynamics are now contained within the complex-valued matrices, which can be analysed using a plethora of well-established numerical methods.
